---
title: "09_Reef_Flat_Crest_and_Forereef_Models"
author: "Stephan Bitterwolf"
date: "2023-07-27"
output:
  html_document:
    toc: true
    number_sections: true
    toc_float: true
    theme: cerulean
    df_print: paged
    code_folding: "hide"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo=FALSE}
pre {
  max-height: 100px;
  overflow-y: auto;
}

pre[class] {
  max-height: 100px;
}

.btn {
    border-width: 0px 0px 0px;
    font-weight: normal;
    text-transform: ;
}
.btn-default {
    color: #2ecc71;
    background-color: #ffffff;
    border-color: #ffffff;
}


.header-section-number::after {
  content: ".";
}
```
# Overview

The purpose of this code is to compare the Ordinary Least Squares Regression for the Hawaiian Island Erosion and Reef dataset with and without autocorrelation correction.

```{r Libraries, message=FALSE, warning=FALSE}
# library(tidyverse)
# library(spatial)

#library a bunch of packages we may (or may not) use - install them first if not installed already. 
library(tidyverse)
library(sf)
library(tmap)
library(here)

```
# Importing and Merging Data
```{r Import Data}
Reefs_and_Erosion <- st_read(here("2_output","modified_shapefiles", "joined_databases","reefs_and_erosion.shp"))
Reefs_and_Erosion$Trns_ID<-as.numeric(Reefs_and_Erosion$Trns_ID)
Reef_Summary <- read.csv(here("2_output","dataframes", "Reef_Summary.csv"))

Reefs_and_Erosion_w_Geometry<-inner_join(Reefs_and_Erosion, Reef_Summary, by= c("Trns_ID"="Uniq_ID"))


Habitat_Merged<-st_read(here("2_output","modified_shapefiles", "habitat","habitat_merged.shp"))
table(st_is_valid(Habitat_Merged))
Habitat_Merged<-st_make_valid(Habitat_Merged)
table(st_is_valid(Habitat_Merged))

Habitat_Merged<- Habitat_Merged %>%
  filter(Zone =="Land" | Zone =="Fore_Reef" | Zone =="Reef_Crest" | Zone == "Reef_Flat")%>%
  mutate(Zone = factor(Zone, levels = c("Land", "Reef_Flat", "Reef_Crest", "Fore_Reef")))

# library(tmap)
# tmap_mode("view")
# tmap_options(check.and.fix = TRUE)
# # tm_basemap("Stamen.Watercolor")+
# tm_basemap(leaflet::providers$Stamen.Toner)+
#   tm_shape(Habitat_Merged)+
#   tm_polygons(col="Zone", palette=c(Land='green', Reef_Flat='white', Reef_Crest='yellow',Fore_Reef='red'))+
#   tm_shape(Reefs_and_Erosion_w_Geometry)+
#   tm_dots(col="SRate_m", shape = 19, border.col=NULL, alpha=0.8, size=.1)

```
# Modifying imported data
1. Calculate new values based on changes between scenarios
2. Filter out nonsensical results
```{r Modifying Data, message=FALSE, warning=FALSE}
###Transform data such that a log or Sqrt transformation can be applied to the dependent variable
library(tidyverse)
Reefs_and_Erosion_w_Geometry$island<- as_factor(Reefs_and_Erosion_w_Geometry$island)
#Reefs_and_Erosion_w_Geometry <- Reefs_and_Erosion_w_Geometry %>%
  #mutate(Island = fct_relevel(island, levels = "Kauai", "Oahu", "Maui")) #Relevel so that Maui comes first (Moving from East to West)
# df_transformed <- df_sbst
# df_transformed$SRate_log <- log(df_transformed$SRate_m + abs(min(df_transformed$SRate_m))+1)
# df_transformed <- df_transformed %>% dplyr::select(-SRate_m) %>% relocate(SRate_log, .before = WEin1)

#Create dataframe for Erosion and Reef Data
Reefs_and_Erosion_w_Geometry <- dplyr::rename(Reefs_and_Erosion_w_Geometry, WE_onspChg = WE_nsCh)
Reefs_and_Erosion_w_Geometry <- Reefs_and_Erosion_w_Geometry %>% distinct(., ero_ID, .keep_all = TRUE)
Reefs_and_Erosion_w_Geometry <- Reefs_and_Erosion_w_Geometry %>% drop_na(., ero_ID)

#Are there any duplicated values
table(duplicated(Reefs_and_Erosion_w_Geometry$ero_ID))

#Add Reef Width and Depth Summary Columns
Reefs_and_Erosion_w_Geometry<-Reefs_and_Erosion_w_Geometry%>%
  group_by(Trns_ID)%>%
  mutate(Reef_Width = rowSums(pick(Zone_Width_Fore_Reef:Zone_Width_Reef_Flat), na.rm = T),
         Min_Reef_Depth = min(pick(Min_Depth_Fore_Reef:Min_Depth_Reef_Crest), na.rm = T ))%>%
  ungroup()

#Add columns to characterize change between reef scenarios
Reefs_and_Erosion_w_Geometry <- Reefs_and_Erosion_w_Geometry %>% mutate(
                    Ru_Chng = Ru2-Ru1,
                    Ru_pChng = (Ru2-Ru1)/Ru1*100,
                    ZS_Chng = ZS2-ZS1,
                    ZS_pChng = (ZS2-ZS1)/ZS1*100,
                    TWL_Chng = TWL2-TWL1,
                    TWL_pChng = (TWL2-TWL1)/TWL1*100,
                    Lf_Chng = Lf2-Lf1,
                    Lf_pChng = (Lf2-Lf1)/Lf1*100,
                    WEin_Chng = WEin2-WEin1,
                    WEin_pChng = (WEin2-WEin1)/WEin1*100,
                    WEout_Chng = WEout2-WEout1,
                    WEout_pChng = (WEout2-WEout1)/WEout1*100
                    )
#Relocate columns
Reefs_and_Erosion_w_Geometry <- Reefs_and_Erosion_w_Geometry %>% relocate(Ru_Chng, .after = Ru2)
Reefs_and_Erosion_w_Geometry <- Reefs_and_Erosion_w_Geometry %>% relocate(Ru_pChng, .after = Ru_Chng)
Reefs_and_Erosion_w_Geometry <- Reefs_and_Erosion_w_Geometry %>% relocate(ZS_Chng, .after = ZS2)
Reefs_and_Erosion_w_Geometry <- Reefs_and_Erosion_w_Geometry %>% relocate(ZS_pChng, .after = ZS_Chng)
Reefs_and_Erosion_w_Geometry <- Reefs_and_Erosion_w_Geometry %>% relocate(TWL_Chng, .after = TWL2)
Reefs_and_Erosion_w_Geometry <- Reefs_and_Erosion_w_Geometry %>% relocate(TWL_pChng, .after = TWL_Chng)
Reefs_and_Erosion_w_Geometry <- Reefs_and_Erosion_w_Geometry %>% relocate(Lf_Chng, .after = Lf2)
Reefs_and_Erosion_w_Geometry <- Reefs_and_Erosion_w_Geometry %>% relocate(Lf_pChng, .after = Lf_Chng)
Reefs_and_Erosion_w_Geometry <- Reefs_and_Erosion_w_Geometry %>% relocate(WEin_Chng, .after = WEin2)
Reefs_and_Erosion_w_Geometry <- Reefs_and_Erosion_w_Geometry %>% relocate(WEin_pChng, .after = WEin_Chng)
Reefs_and_Erosion_w_Geometry <- Reefs_and_Erosion_w_Geometry %>% relocate(WEout_Chng, .after = WEout2)
Reefs_and_Erosion_w_Geometry <- Reefs_and_Erosion_w_Geometry %>% relocate(WEout_pChng, .after = WEout_Chng)



df <- Reefs_and_Erosion_w_Geometry
# df$Cluster<- as.numeric(df$Cluster)
df_sbst <- df %>% 
  dplyr::filter(flag == "Good") %>% #exclude values where reefs are better at reducing wave energy AFTER 1 m reef height was lost
  dplyr::filter(Lr2c1 > -2) #%>% #exclude values where the reef is "onshore" a distance greater than 2 m
  # dplyr::select(SRate_m, Ru1,  Ru_pChng, ZS1,  ZS_pChng, TWL1,  TWL_pChng, Lf1, Lf_pChng, WEin1, WEout1, WEout_pChng, hrf0, hrfmin, Lr2c1, hr2c1, Lreef1, WEred, TWLred, Zsred, WE_onshr1, WE_onspChg, Island)

#Add Unique Island and AreaName Column
df_sbst$Islxnam <- paste(df_sbst$Island," @ ", df_sbst$AreaNam, sep="")

#Move String Columns to End of Dataframe
df_sbst <- df_sbst %>% relocate(c(Island, AreaNam, Islxnam), .after = last_col())

#Calculate fold change
df_sbst<-df_sbst%>%
mutate(logit_WE_onshr1 = car::logit(WE_nsh1), 
         fold_WE_onsh_Chg = (WE_onspChg/100), 
         fold_Ru_Chng = (Ru_pChng/100),
         fold_ZS_Chng = (ZS_pChng/100),
         fold_TWL_Chng = (TWL_pChng/100),
         fold_Lf_Chng = (Lf_pChng/100),
         #fold_WEin_Chng = (WEin_pChng/100),
         fold_WEout_Chng = (WEout_pChng/100),
         .keep = "unused")

#Move new columns
df_sbst <- df_sbst %>% relocate(fold_Ru_Chng, .after = Ru1)
df_sbst <- df_sbst %>% relocate(fold_ZS_Chng, .after = ZS1)
df_sbst <- df_sbst %>% relocate(fold_TWL_Chng, .after = TWL1)
df_sbst <- df_sbst %>% relocate(fold_Lf_Chng, .after = Lf1)
#df_transformed <- df_transformed %>% relocate(fold_WEin_Chng, .after = WEin_Chng)

###Remove Variables Highly Correlated with One Another
data <- df_sbst %>% dplyr::select(-c(Zsred, WEred, TWLred, hrf0))
glimpse(data)

data<- data%>%
  dplyr::select(-c(ID, OBJECTI,FID_,X,Refs_ID, trns_ID_1, Ersn_ID,Island_1,island))



write_csv(data, here("2_output", "dataframes","reefs_and_erosion_w_geometry.csv"))
st_write(data, here("2_output", "modified_shapefiles", "joined_databases","reefs_and_erosion_w_geometry.gpkg"), delete_layer = TRUE)
```


# Reef Flat Regressions

## Select and Aggregate Variables for the Regression
```{r Select Data, warning=FALSE}


#Select Reef Data
##### Reef Flat #####
kept_vars_F <- c("reef_ID",
               "SRate_m",         #Shoreline Change rate (m/yr)
               "Sncrt_m", #uncertainty in the change rate (m/yr) used as part of the weighted average
               
               #"Island", ##Removed due to conversation with Mike in Jan of 2023
               #  "Maui",
               # "Kauai",
               # "Oahu",

               ### MODEL FORCING ###
               "WEin1",             #Wave energy entering the model @ 30 m depth

               ### NEW REEF GEOMETRY ###
                # "Distance_Offshore_Fore_Reef",
                # "Distance_Offshore_Reef_Crest",
                "Distance_Offshore_Reef_Flat",
                # "Reef_Width",
                # # # "Zone_Width_Fore_Reef",
                # # "Zone_Width_Reef_Crest",
                "Zone_Width_Reef_Flat",
                # # # "Median_Depth_Fore_Reef",
                # # "Median_Depth_Reef_Crest",
                "Median_Depth_Reef_Flat",
                # # # # "Max_Depth_Fore_Reef",
                # # # "Max_Depth_Reef_Crest",
                "Max_Depth_Reef_Flat",
                # # # # "Min_Depth_Fore_Reef",
                # # # "Min_Depth_Reef_Crest",
                 "Min_Depth_Reef_Flat",
               # "Min_Reef_Depth"
                # # # "slope_Fore_Reef",
                # # "slope_Reef_Crest",
                 "slope_Reef_Flat"

               # ### Model Output or Calculations ###
               # "TWL1",              #Total water Level At Shore (i.e., model output location) ##Removed in favor of Ru1 due to conversation with Mike in Jan of 2023
               # "logit_WE_onshr1",   # Proportion of incoming wave energy arriving onshore (logit transformed)
               # "Ru1",
               # # 
               # # ### Reef Degradation Impact ###
               # # #these variables calculate the relative impact of reef deepening by 1 m on model outputs
               # "fold_TWL_Chng",       #Fold change in total water level at shoreline between current and degraded reefs bathymetry (-1 m)
               # "fold_WE_onsh_Chg",     #Fold change in wave energy at shoreline between current and degraded reefs bathymetry (-1 m)
               # "fold_Lf_Chng"
               ) #Here I write which variables to include in the model
##### Reef Flat + Crest #####
kept_vars_FC <- c("reef_ID",
               "SRate_m",         #Shoreline Change rate (m/yr)
               "Sncrt_m", #uncertainty in the change rate (m/yr) used as part of the weighted average
               
               #"Island", ##Removed due to conversation with Mike in Jan of 2023
               #  "Maui",
               # "Kauai",
               # "Oahu",

               ### MODEL FORCING ###
               "WEin1",             #Wave energy entering the model @ 30 m depth

               ### NEW REEF GEOMETRY ###
                # "Distance_Offshore_Fore_Reef",
                "Distance_Offshore_Reef_Crest",
                "Distance_Offshore_Reef_Flat",
                # "Reef_Width",
                # # # "Zone_Width_Fore_Reef",
                "Zone_Width_Reef_Crest",
                "Zone_Width_Reef_Flat",
                # # # "Median_Depth_Fore_Reef",
                "Median_Depth_Reef_Crest",
                "Median_Depth_Reef_Flat",
                # # # # "Max_Depth_Fore_Reef",
                "Max_Depth_Reef_Crest",
                "Max_Depth_Reef_Flat",
                # # # # "Min_Depth_Fore_Reef",
                "Min_Depth_Reef_Crest",
                 "Min_Depth_Reef_Flat",
               "Min_Reef_Depth",
                # # # "slope_Fore_Reef",
                "slope_Reef_Crest",
                 "slope_Reef_Flat"

               # ### Model Output or Calculations ###
               # "TWL1",              #Total water Level At Shore (i.e., model output location) ##Removed in favor of Ru1 due to conversation with Mike in Jan of 2023
               # "logit_WE_onshr1",   # Proportion of incoming wave energy arriving onshore (logit transformed)
               # "Ru1",
               # # 
               # # ### Reef Degradation Impact ###
               # # #these variables calculate the relative impact of reef deepening by 1 m on model outputs
               # "fold_TWL_Chng",       #Fold change in total water level at shoreline between current and degraded reefs bathymetry (-1 m)
               # "fold_WE_onsh_Chg",     #Fold change in wave energy at shoreline between current and degraded reefs bathymetry (-1 m)
               # "fold_Lf_Chng"
               ) #Here I write which variables to include in the model

##### Reef Flat + Crest + Forereef#####
kept_vars_FCF <- c("reef_ID",
               "SRate_m",         #Shoreline Change rate (m/yr)
               "Sncrt_m", #uncertainty in the change rate (m/yr) used as part of the weighted average
               
               #"Island", ##Removed due to conversation with Mike in Jan of 2023
               #  "Maui",
               # "Kauai",
               # "Oahu",

               ### MODEL FORCING ###
               "WEin1",             #Wave energy entering the model @ 30 m depth

               ### NEW REEF GEOMETRY ###
                "Distance_Offshore_Fore_Reef",
                "Distance_Offshore_Reef_Crest",
                "Distance_Offshore_Reef_Flat",
                "Reef_Width",
                "Zone_Width_Fore_Reef",
                "Zone_Width_Reef_Crest",
                "Zone_Width_Reef_Flat",
                "Median_Depth_Fore_Reef",
                "Median_Depth_Reef_Crest",
                "Median_Depth_Reef_Flat",
                "Max_Depth_Fore_Reef",
                "Max_Depth_Reef_Crest",
                "Max_Depth_Reef_Flat",
                "Min_Depth_Fore_Reef",
                "Min_Depth_Reef_Crest",
                 "Min_Depth_Reef_Flat",
               "Min_Reef_Depth",
                "slope_Fore_Reef",
                "slope_Reef_Crest",
                 "slope_Reef_Flat"

               # ### Model Output or Calculations ###
               # "TWL1",              #Total water Level At Shore (i.e., model output location) ##Removed in favor of Ru1 due to conversation with Mike in Jan of 2023
               # "logit_WE_onshr1",   # Proportion of incoming wave energy arriving onshore (logit transformed)
               # "Ru1",
               # # 
               # # ### Reef Degradation Impact ###
               # # #these variables calculate the relative impact of reef deepening by 1 m on model outputs
               # "fold_TWL_Chng",       #Fold change in total water level at shoreline between current and degraded reefs bathymetry (-1 m)
               # "fold_WE_onsh_Chg",     #Fold change in wave energy at shoreline between current and degraded reefs bathymetry (-1 m)
               # "fold_Lf_Chng"
               ) #Here I write which variables to include in the model

##### Aggregating #####
#Save only important data

## Non Aggregated Dataset
Reef_data <- data %>%
  #select(-AreaName,-X,-Uniq_ID,-...1, -Transect_Type,-Erosion_Data, -SRate_log)%>%
  dplyr::select(all_of(kept_vars_FCF))%>%
  na.omit(.)

## Detect outliers on for each reef_ID

weighted_mean <- function(rate, uncertainty) {
  weights <- 1 / uncertainty  # Inverse of uncertainties as weights
  weighted_mean <- sum(rate * weights) / sum(weights)
  return(weighted_mean)
}

### Use Tukey's fences to remove outliers but keep sites where there is only 1 erosion value
data_cleaned <- Reef_data %>%
  group_by(reef_ID, geometry) %>%
  mutate(
    Q1 = quantile(SRate_m, 0.25),
    Q3 = quantile(SRate_m, 0.75),
    IQR = Q3 - Q1,
    Lower_Fence = Q1 - 1.5 * IQR,
    Upper_Fence = Q3 + 1.5 * IQR,
    n_erosion=n()
  ) %>%
  mutate(outlier=if_else(n_erosion==1,FALSE,(SRate_m <= Lower_Fence & SRate_m >= Upper_Fence)))%>% #Keep sites where there is only one erosion value
  mutate(n_outlier=sum(outlier))%>%
  ungroup()

### Note, there are no outliers in this dataset on a site by site basis.

## Aggregated Dataset using the median and the weighted mean **Note these aggregations will only affect the SRate_m value
weighted_mean <- 
  data_cleaned%>%
  group_by(reef_ID, geometry)%>%
  mutate(transformed_uncertainty= 1/Sncrt_m) %>%
  summarise(wgt_mean_SRate=weighted.mean(SRate_m,transformed_uncertainty))

summarised_dataset<-Reef_data%>%
  group_by(reef_ID, geometry)%>%
  mutate(transformed_uncertainty= 1/Sncrt_m) %>%
  summarise_all(median)

Reef_data_aggregated<-st_join(weighted_mean,summarised_dataset, by="reef_ID")%>%
  select(-c(reef_ID.y, transformed_uncertainty,Sncrt_m))%>%
  rename(reef_ID=reef_ID.x,
         median_SRate_m=SRate_m)

class(Reef_data_aggregated)
st_write(Reef_data_aggregated, here("2_output", "modified_shapefiles", "joined_databases", "Reef_data_aggregated.gpkg"), delete_layer = TRUE)

```

```{r Plot Some Aggregation Comparisons}
Reef_data_aggregated%>%
  ggplot()+
  geom_point(aes(x=as_factor(reef_ID),y=wgt_mean_SRate, color="Weighted Mean"))+
  geom_point(aes(x=as_factor(reef_ID),y=median_SRate_m, color="Median"))+
  scale_color_manual(values = c("Weighted Mean" = "blue", "Median" = "red"))+
  labs(y="Erosion Rate",
       x="Site",
       title="Comparing Erosion Rate Summarization Methods")

Reef_data_aggregated%>%
  ggplot()+
  geom_point(aes(x=as_factor(reef_ID),y=wgt_mean_SRate-median_SRate_m))+
  labs(y="Erosion Rate",
       x="Site",
       title="Difference between weighted mean and median")


Reef_data_aggregated%>%
  ggplot()+
  geom_point(aes(x=WEin1,y=wgt_mean_SRate, color="Weighted Mean"))+
  geom_point(aes(x=WEin1,y=median_SRate_m, color="Median"))+
  labs(y="Erosion Rate",
       x="Site",
       title=" Correlation between WEin1 and the two Erosion Rate methods")


```
The above plots show that the Weighted mean method seems fine. Since there is large variability in uncertainty, I would like to bias the data towards lower uncertainties.

```{r Select Summarization Method}
#Here I will choose to use the weighted mean instead of the median

Reef_data_aggregated <- Reef_data_aggregated %>%
  select(-c(median_SRate_m, reef_ID))

```



## Create models with Non-Transformed Data

### Linear Model

```{r Plot Linear Relationships}
dataset<-Reef_data_aggregated

dependent_variable<-"wgt_mean_SRate"

plot(st_drop_geometry(dataset))

library(corrplot)
corplot.data<-st_drop_geometry(dataset) %>%
  select(where(is.numeric)) %>%
  ungroup()

xtr = corplot.data %>% dplyr::select(-dependent_variable)
ytr = corplot.data %>% dplyr::select(dependent_variable)

cor_numVar = cor(cbind(xtr, ytr), use="pairwise.complete.obs") # correlations of all numeric variables
# sort on decreasing correlations with log1p_SalePrice      
cor_sorted = as.matrix(sort(cor_numVar[, dependent_variable], decreasing = TRUE))
  
# select only high corelations
CorHigh = names(which(apply(cor_sorted, 1, function(x) abs(x)>0.1)))
cor_numVar = cor_numVar[CorHigh, CorHigh]

corrplot.mixed(cor_numVar, tl.col="black", tl.pos = "lt")

## Plot Correlated Variables
dataset %>%
  ggplot(aes(y=Reef_Width, x=Zone_Width_Reef_Flat))+
  geom_point()+
  stat_smooth()


```


### Variable Selection with LASSO



```{r LASSO Selection 1}
model_data <-st_drop_geometry(dataset)

#Here I follow the example from Julia SIlge: https://juliasilge.com/blog/lasso-the-office/
library(tidymodels)
erosion_split <- initial_split(model_data, prop=3/4, strata = dependent_variable)

erosion_train <- training(erosion_split)
erosion_train %>% glimpse()

erosion_test <- testing(erosion_split)
erosion_test %>% glimpse()

erosion_rec <- recipe(as.formula(paste(dependent_variable, "~ .")), data = erosion_train) %>%
    step_zv(all_numeric(), -all_outcomes()) %>%
    step_dummy(all_nominal(), one_hot = TRUE)%>%
    # step_BoxCox(all_numeric(), -all_outcomes())%>%#We are scaling and standardizing the dummies as well as the numeric datahttps://stats.stackexchange.com/questions/69568/whether-to-rescale-indicator-binary-dummy-predictors-for-lasso
    step_normalize(all_numeric(), -all_outcomes())

erosion_prep <- erosion_rec

###

lasso_spec <- linear_reg(penalty = 0.1, mixture = 1) %>%
  set_engine("glmnet")

wf <- workflow() %>%
  add_recipe(erosion_rec)

lasso_fit <- wf %>%
  add_model(lasso_spec) %>%
  fit(data = erosion_train)

lasso_fit %>%
  pull_workflow_fit() %>%
  tidy()


###

set.seed(1234)
erosion_boot <- bootstraps(erosion_train, strata = dependent_variable)

tune_spec <- linear_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

lambda_grid <- grid_regular(penalty(), levels = 100)



###
doParallel::registerDoParallel()

set.seed(2020)
lasso_grid <- tune_grid(
  wf %>% add_model(tune_spec),
  resamples = erosion_boot,
  grid = lambda_grid
)

###

lasso_grid %>%
  collect_metrics()


###

lasso_grid %>%
  collect_metrics() %>%
  ggplot(aes(penalty, mean, color = .metric)) +
  geom_errorbar(aes(
    ymin = mean - std_err,
    ymax = mean + std_err
  ),
  alpha = 0.5
  ) +
  geom_line(size = 1.5) +
  facet_wrap(~.metric, scales = "free", nrow = 2) +
  scale_x_log10() +
  theme(legend.position = "none")

###

lowest_rmse <- lasso_grid %>%
  select_best("rmse")

final_lasso <- finalize_workflow(
  wf %>% add_model(tune_spec),
  lowest_rmse
)

###

library(vip)

final_lasso %>%
  fit(erosion_train) %>%
  pull_workflow_fit() %>%
  vi(lambda = lowest_rmse$penalty) %>%
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, Importance)
  ) %>%
  ggplot(aes(x = Importance, y = Variable, fill = Sign)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)


last_fit(
  final_lasso,
  erosion_split
) %>%
  collect_metrics()

```
The lasso model selects for 5 out of the 7 variables. slope_Reef_Flat and Median_Depth_Reef_Flat are omitted.

```{r Lasso Omitted Values}

lasso_rejected_explanatory_variables<- final_lasso %>%
  fit(erosion_train) %>%
  pull_workflow_fit() %>%
  vi(lambda = lowest_rmse$penalty) %>%
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, Importance))%>%
  filter(Importance==0)


lasso_selected_dataset <- dataset[, !colnames(dataset) %in% lasso_rejected_explanatory_variables$Variable]



```

```{r Build Model No Transformation}
dataset<-lasso_selected_dataset

library(tidymodels)
model_data<-st_drop_geometry(dataset) %>%
   ungroup()
  

model_recipe <-
    recipe(as.formula(paste(dependent_variable, "~ .")), data=model_data)%>%
  step_center(all_numeric_predictors())%>%
  step_scale(all_numeric_predictors())%>%
  prep

lm_spec <- linear_reg() %>% set_engine("lm")


lm_fit <- fit(lm_spec, as.formula(paste(dependent_variable, "~ .")), data=bake(model_recipe, new_data = NULL))

regression_data<- bake(model_recipe, new_data = NULL)

tidy(lm_fit)
summary(lm_fit$fit)


```


#### Plot Model Results
```{r Plot Model Results}
#install.packages('scico')
library(scico)

tidy(lm_fit)%>%
  ggplot(aes(x=term, y=estimate, fill=p.value, color=p.value))+
  geom_point()+
  geom_bar(stat="identity")+
  scale_fill_scico("p",palette = 'vik')+
  scale_color_scico("p",palette = 'vik')+
  coord_flip()+
  labs(x="",y="estimate")


#Plot Residuals
hist(lm_fit$fit$residuals)

{par(mfrow=c(2,2)) # plot all 4 plots in one

plot(lm_fit$fit, 
     pch = 16,    # optional parameters to make points blue
     col = '#006EA1')
}



```


#### Bootstrap Resampling of Linear Model

```{r BootStrap Simple Regression}
library(tidymodels)


set.seed(1992)

#GET CENTERED AND SCALED DATA for Regression
data_boot <- model_data%>%
    recipe(as.formula(paste(dependent_variable, "~ .")))%>%
    step_center(all_numeric_predictors())%>%
    step_scale(all_numeric_predictors())%>%
    prep %>% 
    bake(NULL)
###
  fit_lm_on_bootstrap <- function(split) {
  linear_reg() %>%
  set_engine("lm")%>%
  fit(SRate_m ~., data=bake(model_recipe, new_data = NULL))
}


#Bootstrap Resample the centered and scaled data
bootstrap_data<- bootstraps(data_boot, times=1000, apparent = TRUE)

erosion_models <- bootstrap_data %>%
  mutate(
    model = map(splits, ~ lm(as.formula(paste(dependent_variable, "~ .")), data = .)),
    coef_info = map(model, tidy))
  



erosion_coefs <- erosion_models %>%
  unnest(coef_info)

#glimpse(erosion_coefs)


erosion_coefs <- erosion_coefs %>% filter(term != "(Intercept)")
#erosion_coefs
erosion_coefs<-as_tibble(erosion_coefs) %>%
  select(-c(splits,model))


library(rsample)
conf_int<-int_pctl(erosion_models, coef_info, alpha = 0.05)

conf_int<-conf_int%>% filter(term != "(Intercept)")
conf_int<-as_tibble(conf_int)

test<-left_join(erosion_coefs,conf_int, by="term")
test <- test %>%
  mutate(sig=if_else(p.value>0.05,">0.05","<0.05"))

test %>%
  ggplot(aes(x=estimate)) +
  geom_histogram(position="identity", alpha=0.5, bins=15)+
  labs(title="Bootstrap resample estimates with 95% confidence intervals",
       x="Coefficient estimates",
       y="Frequency")+
    facet_wrap(~term, scales = "free")+
  #shade_confidence_interval(endpoints = c(test$.lower,test$.upper), fill="gold", alpha=0.1, color=)
  geom_vline(aes(xintercept=.lower))+
  geom_vline(aes(xintercept=.upper))+
  labs(title="Bootstrap resample estimates with 95% confidence intervals",
       x="Coefficient estimates",
       y="Frequency")+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))

### Plotting Bootstrap Coefficients
library(scico)
test%>%
  ggplot(aes(x=term, y=estimate))+
  geom_point(alpha=0.1, position=position_jitter(height=0, width=0.05), aes(color=sig)) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3)+ 
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", 
               width=1, colour="red", alpha=0.7) +
  scale_color_viridis_d("p value",guide = guide_legend(override.aes = list(size = 3,
                                                                    alpha = 1) ))+
  coord_flip()+
  labs(caption="Fig x. Coefficient estimates from 1000 bootstrap resamples. Red dot = mean. Errorbars = 95% Confidence intervals")


```

### Spatial Model


```{r Examine Autocorrelation, message=FALSE, warning=FALSE}
library(sf)
library(sfdep)
library(dplyr)

# grab geometry
geo <- st_geometry(dataset) 

# grab transformed and scaled data
geo_data<-st_drop_geometry(dataset)%>%
    recipe(as.formula(paste(dependent_variable, "~ .")))%>%
    step_center(all_numeric_predictors())%>%
    step_scale(all_numeric_predictors())%>%
    prep %>% 
    bake(NULL)
#join geometry with data

 geo_data_map<- dataset

#save model residuals to our dataset
geo_data_map$residuals<-lm_fit$fit$residuals

#save file for external use
st_write(geo_data_map,"geo_data_map.shp", append=FALSE)

#### Calculating Neighbors ####
#Calculate neighbors based on a distance band
critical_threshold(st_geometry(geo_data_map))
nb_dist <- st_dist_band(st_geometry(geo_data_map),  upper = critical_threshold(st_geometry(geo_data_map))) 

#Calculate neighbors based on kneigherst neighbors
nb_knear_8<- st_knn(geometry=st_geometry(geo_data_map), k = 8)


#### Calculate Moran's I ####
#*# Decide which neighbors to use
neighbors <-nb_dist 
# neighbors <-nb_knear_8 
#Create weights matrix
wt <- st_weights(neighbors)

#Compute the Moran I and Test to see if it is different from a random distribution
moran<-sfdep::global_moran_test(geo_data_map$residuals,neighbors,wt)
moran

#Re-compute the Moran with a monte carlo simulation
MC<-spdep::moran.mc(geo_data_map$residuals,spdep::nb2listw(neighbors,style = "C"), nsim=599)
MC
plot(MC,main="", las=1)

#Extract Moran Data for Plotting
moran_data<-spdep::moran.plot(geo_data_map$residuals,spdep::nb2listw(neighbors,style = "C"), return_df = TRUE, plot=FALSE)
#trace(ggpubr:::.stat_lm, edit = TRUE) #https://stackoverflow.com/questions/66177005/im-using-stat-regline-equation-with-ggscatter-is-there-a-way-to-specify-the-si
moran_data<-cbind(moran_data, geo_data)

#Plot Moran Data
moran_data %>%
  ggplot(aes(x=x, y=wx, color=get(dependent_variable)))+
  geom_point()+
  labs(x="Model Residuals", y="Spatially Lagged Residuals", caption=paste("Moran's I: ",round(moran$estimate[1],3)))+
  geom_smooth(method= 'lm',formula = y~x, se = FALSE)+
  geom_hline(yintercept = 0, linetype="dashed")+
  geom_vline(xintercept = 0, linetype="dashed")+
  scico::scale_color_scico("Shoreline Change Rate (m/yr)",midpoint = 0, palette = "roma")+
  theme_dark()
  # ggpubr::stat_regline_equation(mapping = NULL,
  #                               data = NULL,
  #                               formula = y ~ x,
  #                               label.x.npc = "left",
  #                               label.y.npc = "top",
  #                               label.x = NULL,
  #                               label.y = NULL,
  #                               output.type = "expression",
  #                               geom = "text",
  #                               position = "identity",
  #                               na.rm = FALSE,
  #                               show.legend = NA,
  #                               inherit.aes = TRUE)

```

The Morans I value is 0.28 meaning that there is spatial autocorrelation between the residuals.

#### Addressing Spatial Autocorrelation
  
```{r Spatial Regressions, include=FALSE}
library(spdep)
library(spatialreg)
#Neighbors= neighbors
nb_list <- nb2listw(neighbors)
#Weights =wt
wt <- wt

### Define Regression Equation ###
lm_fit[["fit"]][["terms"]]
reg.eq= as.formula(paste(dependent_variable, "~ ."))

### Linear Model: OLS ###
linear_model<-lm(reg.eq, data=data_boot)
summary(linear_model)

### SLX ###
SLX_model <- spatialreg::lmSLX(reg.eq, data=data_boot, nb_list) #Does not work for K-nearest neighbors
summary(SLX_model)
impacts(SLX_model, nb_list) #Tells you how neighboring values impact the effects
summary(impacts(SLX_model, nb_list, R=500), zstats=TRUE)#tells you if the effects are significant

### SAR: Spatial Lag Model ###
SAR_model <- lagsarlm(reg.eq, data=data_boot, nb_list)
summary(SAR_model)
impacts(SAR_model,listw=nb_list)
summary(impacts(SAR_model,listw=nb_list, R=2000), zstats=TRUE)

### SEM: Spatial Error Model ###
SEM_model <-errorsarlm(reg.eq, data=data_boot, nb_list)
summary(SEM_model)
Hausman.test(SEM_model) # The high p value says that we cannot reject that the SEM_model is inappropriate

### SDEM: Spatial  Durbin Error Model
SDEM_model<-errorsarlm(reg.eq, data=data_boot, nb_list, etype="emixed")
summary(SDEM_model)
impacts(SDEM_model,listw=nb_list)
summary(impacts(SDEM_model,listw=nb_list, R=2000), zstats=TRUE)
#Compare with maximul likelihood tests and AIC
AIC(linear_model)
AIC(SLX_model)
AIC(SAR_model)
AIC(SEM_model)
AIC(SDEM_model)

cbind(c("Linear",
        "SLX",
        "SAR",
        "SEM",
        "SDEM"),
      c(AIC(linear_model),
AIC(SLX_model),
AIC(SAR_model),
AIC(SEM_model),
AIC(SDEM_model)))

###

```

Because I think my model is a local model (nearby values affecting nearby but not the entire dataset). I will focus on the local models. The Spatial Durbin Model will be compared to the SEM and the SLX

```{r comparing local spatial models}
summary(SDEM_model)
impacts(SDEM_model,listw=nb_list)
summary(impacts(SDEM_model,listw=nb_list, R=2000), zstats=TRUE)
# None of the lagged values are statistically significant

#LR Tests
LR.Sarlm(SDEM_model,SEM_model)
LR.Sarlm(SDEM_model,SLX_model)
LR.Sarlm(SDEM_model,SAR_model)
LR.Sarlm(SDEM_model,linear_model)


#Also check AIC
AIC(SDEM_model)
AIC(SEM_model)
AIC(SLX_model)
AIC(SAR_model)
AIC(linear_model)




#The Test Indicates that I should Restrict the model to a spatial error model
```

The SDEM model has the lowest AIC value. Likelihood ratio comparisons indicate that it is not very different from the SAR or SEM model. I lean in favor of the SEM model since it is easy to interpret.

```{r SEM model tests}

### Heteroskedasticity
bptest.Sarlm(SEM_model, studentize = TRUE)
Hausman.test(SEM_model)
#bptest.Sarlm(SDEM_model, studentize = TRUE)
#The p value indicates that there is not significant heteroskedasticity @ an alpha of 0.05 and p value of 0.065

### Pseudo R^2: 1-(sum of squared estimate of errors/
1-(SEM_model$SSE/(var(data_boot$wgt_mean_SRate)*(length(data_boot$wgt_mean_SRate)-1)))
```
The above output indicates non-significant heteroskedasticity (BP = 3.7435, df = 5, p-value = 0.5869). The pseudo R squared value is 0.30 which is much better than the linear model.

The SEM model: SEM Spatial Error Model
ð‘¦=ð‘‹ð›½+ð‘¢, ð‘¢=ðœ†ð‘Šð‘¢+ðœ€

```{r final model, eval=FALSE, include=FALSE}
# load package
library(sjPlot)
library(sjmisc)
library(sjlabelled)

summary(SEM_model)

SEM_model_data<-tidy(SEM_model)
SEM_model_data %>%
  ggplot(aes(x=term, y=estimate))+
  geom_point()+
  coord_flip()


#Bootstrap Resample the centered and scaled data
bootstrap_data<- bootstraps(data_boot, times=10, apparent = TRUE)

erosion_models2 <- bootstrap_data %>%
  mutate(
    model = map(splits, ~ errorsarlm(reg.eq, data=., nb_list)),
    coef_info = map(model, tidy))
  


erosion_coefs <- erosion_models2 %>%
  unnest(coef_info)

#glimpse(erosion_coefs)

erosion_coefs <- erosion_coefs %>% filter(term != "(Intercept)")
#erosion_coefs
erosion_coefs<-as_tibble(erosion_coefs) %>%
  select(-c(splits,model))


library(rsample)
conf_int<-int_pctl(erosion_models2, coef_info, alpha = 0.05)

conf_int<-conf_int%>% filter(term != "(Intercept)")
conf_int<-as_tibble(conf_int)

test<-left_join(erosion_coefs,conf_int, by="term")
test <- test %>%
  mutate(sig=if_else(p.value>0.05,">0.05","<0.05"))

test %>%
  ggplot(aes(x=estimate)) +
  geom_histogram(position="identity", alpha=0.5, bins=15)+
  labs(title="Bootstrap resample estimates with 95% confidence intervals",
       x="Coefficient estimates",
       y="Frequency")+
    facet_wrap(~term, scales = "free")+
  #shade_confidence_interval(endpoints = c(test$.lower,test$.upper), fill="gold", alpha=0.1, color=)
  geom_vline(aes(xintercept=.lower))+
  geom_vline(aes(xintercept=.upper))+
  labs(title="Bootstrap resample estimates with 95% confidence intervals",
       x="Coefficient estimates",
       y="Frequency")+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))

### Plotting Bootstrap Coefficients
library(scico)
test%>%
  ggplot(aes(x=term, y=estimate))+
  geom_point(alpha=0.1, position=position_jitter(height=0, width=0.05), aes(color=sig)) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3)+ 
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", 
               width=1, colour="red", alpha=0.7) +
  scale_color_viridis_d("p value",guide = guide_legend(override.aes = list(size = 3,
                                                                    alpha = 1) ))+
  coord_flip()+
  labs(caption="Fig x. Coefficient estimates from 1000 bootstrap resamples. Red dot = mean. Errorbars = 95% Confidence intervals")


```


# Create models with Transformed Data

## Transform Predictors Using BoxCox Method

```{r Transform Predictors}
### SELECT DATASET ###
dataset <- Reef_data_aggregated


### END SELECTION ###

xall <-st_drop_geometry(dataset)%>% 
  dplyr::select(-dependent_variable)
xall[1:ncol(xall)] <- lapply(xall[1:ncol(xall)], as.numeric)
num_names = colnames(xall)[sapply(xall, class) == "numeric"]
xall_before_boxcox = xall

  lambdas <- list()
for (v in num_names){
  # box-cox need all of the values to be positive
  xall[v] = xall[v] - min(xall[v]) + 1 # plus the minimun and 1 to prepare for the box-cox
  # search for the best lambda for the box-cox
  low = -2
  up = 2
  for (i in 1:5){
    bc = MASS::boxcox(xall[v][[1]] ~ 1, lambda=seq(low, up, len=(up-low)*100+1), plotit=FALSE)
    best_lambda = bc$x[which(bc$y == max(bc$y))]
    lambdas[[v]] <- best_lambda
    if (best_lambda == up){
      low = best_lambda - 0.01
      up = 2*up
    }
    else if(best_lambda == low){
      up = best_lambda + 0.01
      low = 2*low
    }
    else{break}
    
    # if (i > 3){print(v)}
  }
  # box-cox transformation
  xall[v] = (xall[v] ^ best_lambda - 1) / best_lambda 
}

xall_after_boxcox = xall

library(moments)
library(gridExtra)


den_skew_plots = function(xall){
  num_names = colnames(xall)[sapply(xall, class) == "numeric"]
  p_li = list()
  
  for (num_var in num_names){
    data = as.data.frame(xall[, num_var])
    colnames(data) = 'temp_name'
    grey_degree = 100 - as.integer(min(60, 10*(abs(skewness(data)))))
        
    p = ggplot(data = data, aes(x = `temp_name`)) + 
      geom_line(stat = 'density', size=1) +
      xlab(paste(num_var, '\n', 'Skew:', round(skewness(data)[[1]], 4))) + 
      theme(panel.background = element_rect(fill = paste0('grey', grey_degree)))
    
    p_li = c(p_li, list(p))
  }
  do.call("grid.arrange", c(p_li, ncol=3))
}

lambdas
den_skew_plots(xall_before_boxcox)
den_skew_plots(xall_after_boxcox)

xall <- xall_after_boxcox ##Save transformed data to xall

dataset<-dataset %>%
  dplyr::select(dependent_variable)%>%
  bind_cols(., xall)

st_write(dataset, here("2_output", "modified_shapefiles", "joined_databases", "simple_regression_transformed.gpkg"),delete_layer = TRUE)
```

```{r Export variations of dataset}

# eroding_only<-dataset %>%
#   filter(SRate_m<0)
# hist(eroding_only$SRate_m)
# hist((eroding_only$SRate_m*-1)^(1/3))  
# st_write(eroding_only, here("2_output", "modified_shapefiles", "joined_databases", "complex_regression-eroding_only.shp"),delete_layer = TRUE)
# eroding_only%>%
#   mutate(tSRate_m=(SRate_m*-1)^(1/3))%>%
#   select(-SRate_m)%>%
#   st_write(here("2_output", "modified_shapefiles", "joined_databases", "complex_regression-eroding_only-cuberoot.shp"),delete_layer = TRUE)

```

## Plot Correlation Matrix
```{r Correlation Matrix, message=FALSE, warning=FALSE}

library(corrplot)
corplot.data<-st_drop_geometry(dataset) %>%
  select(where(is.numeric))

xtr = corplot.data %>% dplyr::select(-dependent_variable)
ytr = corplot.data %>% dplyr::select(dependent_variable)

cor_numVar = cor(cbind(xtr, ytr), use="pairwise.complete.obs") # correlations of all numeric variables
# sort on decreasing correlations with log1p_SalePrice      
cor_sorted = as.matrix(sort(cor_numVar[, dependent_variable], decreasing = TRUE))
  
# select only high corelations
CorHigh = names(which(apply(cor_sorted, 1, function(x) abs(x)>0.1)))
cor_numVar = cor_numVar[CorHigh, CorHigh]

corrplot.mixed(cor_numVar, tl.col="black", tl.pos = "lt")

# ## Plot Correlated Variables
# dataset %>%
#   ggplot(aes(x=TWL1, y=Ru1))+
#   geom_point()+
#   stat_smooth()
# hist(dataset$TWL1)
# hist(dataset$Ru1)

# dataset %>%
#   ggplot(aes(x=Distance_Offshore_Reef_Crest, y=Zone_Width_Reef_Flat))+
#   geom_point()+
#   stat_smooth()

# hist(dataset$Zone_Width_Reef_Flat)
# # hist(dataset$Distance_Offshore_Reef_Crest)
  
```
There is small correlation between reef width and SRate but no other variables.






## Variable Selection with LASSO



```{r LASSO Selection 2}
model_data <-st_drop_geometry(dataset)

#Here I follow the example from Julia SIlge: https://juliasilge.com/blog/lasso-the-office/
library(tidymodels)
erosion_split <- initial_split(model_data, prop=3/4, strata = dependent_variable)

erosion_train <- training(erosion_split)
erosion_train %>% glimpse()

erosion_test <- testing(erosion_split)
erosion_test %>% glimpse()

erosion_rec <- recipe(as.formula(paste(dependent_variable, "~ .")), data = erosion_train) %>%
    step_zv(all_numeric(), -all_outcomes()) %>%
    step_dummy(all_nominal(), one_hot = TRUE)%>%
    # step_BoxCox(all_numeric(), -all_outcomes())%>%#We are scaling and standardizing the dummies as well as the numeric datahttps://stats.stackexchange.com/questions/69568/whether-to-rescale-indicator-binary-dummy-predictors-for-lasso
    step_normalize(all_numeric(), -all_outcomes())

erosion_prep <- erosion_rec

###

lasso_spec <- linear_reg(penalty = 0.1, mixture = 1) %>%
  set_engine("glmnet")

wf <- workflow() %>%
  add_recipe(erosion_rec)

lasso_fit <- wf %>%
  add_model(lasso_spec) %>%
  fit(data = erosion_train)

lasso_fit %>%
  pull_workflow_fit() %>%
  tidy()


###

set.seed(1234)
erosion_boot <- bootstraps(erosion_train, strata = dependent_variable)

tune_spec <- linear_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

lambda_grid <- grid_regular(penalty(), levels = 100)



###
doParallel::registerDoParallel()

set.seed(2020)
lasso_grid <- tune_grid(
  wf %>% add_model(tune_spec),
  resamples = erosion_boot,
  grid = lambda_grid
)

###

lasso_grid %>%
  collect_metrics()


###

lasso_grid %>%
  collect_metrics() %>%
  ggplot(aes(penalty, mean, color = .metric)) +
  geom_errorbar(aes(
    ymin = mean - std_err,
    ymax = mean + std_err
  ),
  alpha = 0.5
  ) +
  geom_line(size = 1.5) +
  facet_wrap(~.metric, scales = "free", nrow = 2) +
  scale_x_log10() +
  theme(legend.position = "none")

###

lowest_rmse <- lasso_grid %>%
  select_best("rmse")

final_lasso <- finalize_workflow(
  wf %>% add_model(tune_spec),
  lowest_rmse
)

###

library(vip)

final_lasso %>%
  fit(erosion_train) %>%
  pull_workflow_fit() %>%
  vi(lambda = lowest_rmse$penalty) %>%
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, Importance)
  ) %>%
  ggplot(aes(x = Importance, y = Variable, fill = Sign)) +
  geom_col() +
  scale_x_continuous(expand = c(0, 0)) +
  labs(y = NULL)


last_fit(
  final_lasso,
  erosion_split
) %>%
  collect_metrics()

```
The lasso model selects for 5 out of the 7 variables. slope_Reef_Flat and Median_Depth_Reef_Flat are omitted.

```{r Lasso Omitted Values 2}

lasso_rejected_explanatory_variables<- final_lasso %>%
  fit(erosion_train) %>%
  pull_workflow_fit() %>%
  vi(lambda = lowest_rmse$penalty) %>%
  mutate(
    Importance = abs(Importance),
    Variable = fct_reorder(Variable, Importance))%>%
  filter(Importance==0)


lasso_selected_dataset <- dataset[, !colnames(dataset) %in% lasso_rejected_explanatory_variables$Variable]



```


## Build OLS model using TidyModels
Here the transformed data is centered around a mean of 0 and scaled so that the standard deviation is = 1. Thereafter a linear model is fit and the output is examined.

```{r Build OLS Model 2}

library(tidymodels)
model_data<-st_drop_geometry(lasso_selected_dataset) %>%
  ungroup()

model_recipe <-
    recipe(reg.eq, data=model_data)%>%
  step_center(all_numeric_predictors())%>%
  step_scale(all_numeric_predictors())%>%
  prep

lm_spec <- linear_reg() %>% set_engine("lm")


lm_fit <- fit(lm_spec, reg.eq, data=bake(model_recipe, new_data = NULL))


tidy(lm_fit)
summary(lm_fit$fit)

```
5 of the 12 variables are significant.

## Plot Model Results
```{r Plot Model Results 2}
#install.packages('scico')
library(scico)

tidy(lm_fit)%>%
  ggplot(aes(x=term, y=p.value, fill=estimate, color=estimate))+
  geom_point()+
  geom_bar(stat="identity")+
  geom_hline(yintercept = 0.05)+
  scale_fill_scico("Coefficient",palette = 'vik', midpoint = 0)+
  scale_color_scico("Coefficient",palette = 'vik', midpoint = 0)+
  coord_flip()+
  labs(x="",y="p")


#Plot Residuals
hist(lm_fit$fit$residuals)

{par(mfrow=c(2,2)) # plot all 4 plots in one

plot(lm_fit$fit, 
     pch = 16,    # optional parameters to make points blue
     col = '#006EA1')
}
```
The vertical line in the above plot indicates a p value of 0.05. We can see that the width of the reef crest, wave energy incoming from the ocean, maximum depth of the reef flat, percentage of wave energy coming to shore, and distance of the reef flat offshore are all significant.

```{r BootStrap Simple Regression 2}
library(tidymodels)


set.seed(1992)

#GET CENTERED AND SCALED DATA for Regression
data_boot <- model_data%>%
    recipe(reg.eq)%>%
    step_center(all_numeric_predictors())%>%
    step_scale(all_numeric_predictors())%>%
    prep %>% 
    bake(NULL)
###
  fit_lm_on_bootstrap <- function(split) {
  linear_reg() %>%
  set_engine("lm")%>%
  fit(reg.eq, data=bake(model_recipe, new_data = NULL))
}


#Bootstrap Resample the centered and scaled data
bootstrap_data<- bootstraps(data_boot, times=1000, apparent = TRUE)

erosion_models <- bootstrap_data %>%
  mutate(
    model = map(splits, ~ lm(reg.eq, data = .)),
    coef_info = map(model, tidy))
  



erosion_coefs <- erosion_models %>%
  unnest(coef_info)

#glimpse(erosion_coefs)


erosion_coefs <- erosion_coefs %>% filter(term != "(Intercept)")
#erosion_coefs
erosion_coefs<-as_tibble(erosion_coefs) %>%
  select(-c(splits,model))


library(rsample)
conf_int<-int_pctl(erosion_models, coef_info, alpha = 0.05)

conf_int<-conf_int%>% filter(term != "(Intercept)")
conf_int<-as_tibble(conf_int)

test<-left_join(erosion_coefs,conf_int, by="term")
test <- test %>%
  mutate(sig=if_else(p.value>0.05,">0.05","<0.05"))

test %>%
  ggplot(aes(x=estimate)) +
  geom_histogram(position="identity", alpha=0.5, bins=15)+
  labs(title="Bootstrap resample estimates with 95% confidence intervals",
       x="Coefficient estimates",
       y="Frequency")+
    facet_wrap(~term, scales = "free")+
  #shade_confidence_interval(endpoints = c(test$.lower,test$.upper), fill="gold", alpha=0.1, color=)
  geom_vline(aes(xintercept=.lower))+
  geom_vline(aes(xintercept=.upper))+
  labs(title="Bootstrap resample estimates with 95% confidence intervals",
       x="Coefficient estimates",
       y="Frequency")+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))

### Plotting Bootstrap Coefficients
library(scico)
test%>%
  ggplot(aes(x=term, y=estimate))+
  geom_point(alpha=0.1, position=position_jitter(height=0, width=0.05), aes(color=sig)) +
  stat_summary(fun=mean, geom="point", fill="red", pch=21, size=3)+ 
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", 
               width=1, colour="red", alpha=0.7) +
  scale_color_viridis_d("p value",guide = guide_legend(override.aes = list(size = 3,
                                                                    alpha = 1) ))+
  coord_flip()+
  labs(caption="Fig x. Coefficient estimates from 1000 bootstrap resamples. Red dot = mean. Errorbars = 95% Confidence intervals")


```

To determine if points are spatially independent from oneanother I will save the residuals and calculate a Moran's I for them. If the Moran's I is more than 0 I will consider this evidence of spatial autocorrelation. Thereafter I will determine which linear model may address this for the final model.




```{r Calculate Morans I 2}
#install.packages("sfdep")

library(sf)
library(sfdep)
library(dplyr)

# grab geometry
geo <- st_geometry(dataset)

# grab transformed and scaled data
geo_data<-st_drop_geometry(dataset)%>%
    recipe(as.formula(paste(dependent_variable, "~ .")))%>%
    # step_center(all_numeric_predictors())%>%
    # step_scale(all_numeric_predictors())%>%
    prep %>% 
    bake(NULL)


#save model residuals to our dataset
geo_data_map$residuals<-lm_fit$fit$residuals

#save file for external use
st_write(geo_data_map,"geo_data_map_transformed_dataset.shp", append=FALSE)

#### Calculating Neighbors ####
#Calculate neighbors based on a distance band
critical_threshold(st_geometry(geo_data_map))
nb_dist <- st_dist_band(st_geometry(geo_data_map),  upper = critical_threshold(st_geometry(geo_data_map))) 

#Calculate neighbors based on kneigherst neighbors
nb_knear_8<- st_knn(geometry=st_geometry(geo_data_map), k = 8)


#### Calculate Moran's I ####
#*# Decide which neighbors to use
neighbors <-nb_dist 

#Create weights matrix
wt <- st_weights(neighbors)

#Compute the Moran I and Test to see if it is different from a random distribution
moran<-sfdep::global_moran_test(geo_data_map$residuals,neighbors,wt)
moran

#Re-compute the Moran with a monte carlo simulation
MC<-spdep::moran.mc(geo_data_map$residuals,spdep::nb2listw(neighbors,style = "C"), nsim=599)
MC
plot(MC,main="", las=1)

#Extract Moran Data for Plotting
moran_data<-spdep::moran.plot(geo_data_map$residuals,spdep::nb2listw(neighbors,style = "C"), return_df = TRUE, plot=FALSE)
#trace(ggpubr:::.stat_lm, edit = TRUE) #https://stackoverflow.com/questions/66177005/im-using-stat-regline-equation-with-ggscatter-is-there-a-way-to-specify-the-si
moran_data<-cbind(moran_data, geo_data)

#Plot Moran Data
moran_data %>%
  ggplot(aes(x=x, y=wx, color=wgt_mean_SRate))+
  geom_point()+
  labs(x="Model Residuals", y="Spatially Lagged Residuals", caption=paste("Moran's I: ",round(moran$estimate[1],3)))+
  geom_smooth(method= 'lm',formula = y~x, se = FALSE)+
  geom_hline(yintercept = 0, linetype="dashed")+
  geom_vline(xintercept = 0, linetype="dashed")+
  scico::scale_color_scico("Shoreline Change Rate (m/yr)",midpoint = 0, palette = "roma")+
  theme_dark()
  # ggpubr::stat_regline_equation(mapping = NULL,
  #                               data = NULL,
  #                               formula = y ~ x,
  #                               label.x.npc = "left",
  #                               label.y.npc = "top",
  #                               label.x = NULL,
  #                               label.y = NULL,
  #                               output.type = "expression",
  #                               geom = "text",
  #                               position = "identity",
  #                               na.rm = FALSE,
  #                               show.legend = NA,
  #                               inherit.aes = TRUE)

```
From the plots we see that our residuals are autocorrelated in space (similar values next to one another). Below I am going to address this by modeling the data using OLS, SLX, SAR, and SEM

```{r Spatial Regressions 2, include=FALSE}
library(spdep)
library(spatialreg)
#Neighbors= neighbors
nb_list <- nb2listw(neighbors)
#Weights =wt
wt <- wt

### Define Regression Equation ###
lm_fit[["fit"]][["terms"]]
reg.eq= as.formula(paste(dependent_variable, "~ ."))

### Linear Model: OLS ###
linear_model<-lm(reg.eq, data=data_boot)
summary(linear_model)

### SLX ###
SLX_model <- spatialreg::lmSLX(reg.eq, data=data_boot, nb_list) #Does not work for K-nearest neighbors
summary(SLX_model)
impacts(SLX_model, nb_list) #Tells you how neighboring values impact the effects
summary(impacts(SLX_model, nb_list, R=500), zstats=TRUE)#tells you if the effects are significant

### SAR: Spatial Lag Model ###
SAR_model <- lagsarlm(reg.eq, data=data_boot, nb_list)
summary(SAR_model)
impacts(SAR_model,listw=nb_list)
summary(impacts(SAR_model,listw=nb_list, R=2000), zstats=TRUE)

### SEM: Spatial Error Model ###
SEM_model <-errorsarlm(reg.eq, data=data_boot, nb_list)
summary(SEM_model)
Hausman.test(SEM_model) # The high p value says that we cannot reject that the SEM_model is inappropriate

### SDEM: Spatial  Durbin Error Model
SDEM_model<-errorsarlm(reg.eq, data=data_boot, nb_list, etype="emixed")
summary(SDEM_model)
impacts(SDEM_model,listw=nb_list)
summary(impacts(SDEM_model,listw=nb_list, R=2000), zstats=TRUE)
#Compare with maximul likelihood tests and AIC
AIC(linear_model)
AIC(SLX_model)
AIC(SAR_model)
AIC(SEM_model)
AIC(SDEM_model)

cbind(c("Linear",
        "SLX",
        "SAR",
        "SEM",
        "SDEM"),
      c(AIC(linear_model),
AIC(SLX_model),
AIC(SAR_model),
AIC(SEM_model),
AIC(SDEM_model)))

###

```

Because I think my model is a local model (nearby values affecting nearby but not the entire dataset). I will focus on the local models. The Spatial Durbin Model will be compared to the SEM and the SLX

```{r comparing local spatial models 2}
summary(SDEM_model)
impacts(SDEM_model,listw=nb_list)
summary(impacts(SDEM_model,listw=nb_list, R=2000), zstats=TRUE)
# None of the lagged values are statistically significant

#LR Tests
LR.Sarlm(SDEM_model,SEM_model)
LR.Sarlm(SDEM_model,SLX_model)
LR.Sarlm(SDEM_model,linear_model)
LR.Sarlm(SDEM_model,SAR_model)


#Also check AIC
AIC(linear_model)
AIC(SLX_model)
AIC(SEM_model)
AIC(SDEM_model)

#The Test Indicates that I should Restrict the model to a spatial error model
```

The spatial DEM model has the lowest AIC and likelihood ratio test indicates that it is better than the spatial durbin error model. Now I will check the SEM model.

```{r SEM model tests 2}

### Heteroskedasticity
bptest.Sarlm(SEM_model, studentize = TRUE)
bptest.Sarlm(SDEM_model, studentize = TRUE)
#The p value indicates that there is not significant heteroskedasticity @ an alpha of 0.05 and p value of 0.065

### Pseudo R^2: 1-(sum of squared estimate of errors/
1-(SEM_model$SSE/(var(data_boot$wgt_mean_SRate)*(length(data_boot$wgt_mean_SRate)-1)))
```
The SEM model: SEM Spatial Error Model
ð‘¦=ð‘‹ð›½+ð‘¢, ð‘¢=ðœ†ð‘Šð‘¢+ðœ€



```{r other things, eval=FALSE, include=FALSE}

g <- geo_data_map %>% 
  mutate(
    nb = st_dist_band(geo,  upper = critical_threshold(geo)),
    wt = st_weights(nb)
    )

library(ggplot2)

# create spatial lag
g %>% 
  mutate(eros_pers_lag = st_lag(SRate_m, nb, wt)) %>% 
  ggplot(aes(color = eros_pers_lag)) + 
  geom_sf(lwd = 0.2) +
  theme_void()


lisa <- g %>% 
  mutate(moran = local_moran(SRate_m, nb, wt))
pull(lisa, moran) %>% 
  glimpse()


lisa %>% 
  tidyr::unnest(moran) %>% 
  mutate(pysal = ifelse(p_folded_sim <= 0.1, as.character(pysal), NA)) |> 
  ggplot(aes(color = pysal)) +
  geom_sf() +
  geom_sf(lwd = 0.2) +
  theme_void() +
  scale_color_manual(values = c("#1C4769", "#24975E", "#EACA97", "#B20016"))
```


